---
title: Cyclomatic complexity
---

```{r cyc-include-setup, include=FALSE}
source(knitr::purl('setup.Rmd', output=tempfile()))
```

This section covers the analysis relating to the _cyclomatic complexity_ metric.

## Data preparation

```{r cyc-data-preparation}
model_seed <- 20230331 # 2023 March 31

create_cyclomatic_complexity_data <- function(data) {
  list(
    cyclomatic_complexity_kloc_std=scale(data$cyclomatic_complexity_kloc),
                  contributors_std=scale(data$contributors),
                         stars_std=scale(data$stars),
                         files_std=scale(data$files),
                          size_std=scale(data$loc),
                           age_std=scale(get_age(data$created_at))
  )
}

data <- create_cyclomatic_complexity_data(df)
data$category <- factor(df$category)
```

## Candidate models {.tabset}

### CYC1

```{r cyc-model-1, message=FALSE}
fit_model_cyc1 <- function(name, data) {
  fit_model(name,
            "cyclomatic_complexity_kloc_std ~ 0 + category",
            c(
              prior(normal(0, 1), class=b),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

cyc1 <- fit_model_cyc1("cyc1", data)
```

### CYC2

```{r cyc-model-2, message=FALSE}
fit_model_cyc2 <- function(name, data) {
  fit_model(name,
            "cyclomatic_complexity_kloc_std ~ 0 + category + contributors_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="contributors_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

cyc2 <- fit_model_cyc2("cyc2", data)
```

### CYC3

```{r cyc-model-3, message=FALSE}
fit_model_cyc3 <- function(name, data) {
  fit_model(name,
            "cyclomatic_complexity_kloc_std ~ 0 + category + size_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="size_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

cyc3 <- fit_model_cyc3("cyc3", data)
```

### CYC4

```{r cyc-model-4, message=FALSE}
fit_model_cyc4 <- function(name, data) {
  fit_model(name,
            "cyclomatic_complexity_kloc_std ~ 0 + category + files_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="files_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

cyc4 <- fit_model_cyc4("cyc4", data)
```

### CYC5

```{r cyc-model-5, message=FALSE}
fit_model_cyc5 <- function(name, data) {
  fit_model(name,
            "cyclomatic_complexity_kloc_std ~ 0 + category + age_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="age_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

cyc5 <- fit_model_cyc5("cyc5", data)
```

### CYC6

```{r cyc-model-6, message=FALSE}
fit_model_cyc6 <- function(name, data) {
  fit_model(name,
            "cyclomatic_complexity_kloc_std ~ 0 + category + stars_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="stars_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

cyc6 <- fit_model_cyc6("cyc6", data)
```

### CYC7

```{r cyc-model-7, message=FALSE}
fit_model_cyc7 <- function(name, data) {
  fit_model(name,
            "cyclomatic_complexity_kloc_std ~ 0 + category + size_std + files_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="size_std"),
              prior(normal(0, 1), class=b, coef="files_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

cyc7 <- fit_model_cyc7("cyc7", data)
```

### CYC8

```{r cyc-model-8, message=FALSE}
fit_model_cyc8 <- function(name, data) {
  fit_model(name,
            "cyclomatic_complexity_kloc_std ~ 0 + category + size_std + age_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="size_std"),
              prior(normal(0, 1), class=b, coef="age_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

cyc8 <- fit_model_cyc8("cyc8", data)
```

## Model comparison

```{r cyc-model-comparison, message=FALSE}
cyc1 <- add_criterion(cyc1, criterion="loo", moment_match=FALSE)
cyc2 <- add_criterion(cyc2, criterion="loo", moment_match=FALSE)
cyc3 <- add_criterion(cyc3, criterion="loo", moment_match=FALSE)
cyc4 <- add_criterion(cyc4, criterion="loo", moment_match=FALSE)
cyc5 <- add_criterion(cyc5, criterion="loo", moment_match=FALSE)
cyc6 <- add_criterion(cyc6, criterion="loo", moment_match=FALSE)
cyc7 <- add_criterion(cyc7, criterion="loo", moment_match=FALSE)
cyc8 <- add_criterion(cyc8, criterion="loo", moment_match=FALSE)

print(loo_compare(cyc1, cyc2, cyc3, cyc4, cyc5, cyc6, cyc7, cyc8, criterion="loo"),
      simplify=FALSE)
```

## Diagnostics {.tabset}

### Prior predictive check

```{r cyc-ppc}
plot_priors(cyc7, "Cyclomatic complexity per KLOC (std)")
```

### MCMC chain trace plots

```{r cyc-caterpillars, message=FALSE}
plot_caterpillar(cyc7, "b_category1", "N")
plot_caterpillar(cyc7, "b_category2", "A")
plot_caterpillar(cyc7, "b_category3", "AE")
plot_caterpillar(cyc7, "b_category5", "AEP")
plot_caterpillar(cyc7, "b_category7", "AEP+")
```

## Full population

### Estimates {.tabset}

#### Standardized

```{r cyc-full-standardized-estimates, message=FALSE}
summary(cyc7)
```

#### Natural scale

```{r cyc-full-natural-scale-estimates, message=FALSE}
print(create_natural_scale_summary_df(cyc7, df$cyclomatic_complexity_kloc))
```

### Plots {.tabset}

#### Major category intervals

```{r cyc-full-major-category-intervals, fig.height=2.8, message=FALSE}
plot_intervals_of_major_categories(cyc7) +
  labs(title="CYC7", x="Cyclomatic complexity per KLOC (std)")
```

#### Binary warning usage

```{r cyc-full-binary-warning-usage, fig.height=2, message=FALSE}
data_binary_factor <- create_cyclomatic_complexity_data(df)
data_binary_factor$category <- factor(uses_warnings(df$category))

cyc7_binary_factor <- fit_model_cyc7("cyc7_factor_uses_warnings",
                                     data_binary_factor)

plot_intervals_of_categories(cyc7_binary_factor) +
  labs(title="CYC7", x="Cyclomatic complexity per KLOC (std)") +
  scale_y_discrete(labels=c(
                     "b_category1"="No warnings",
                     "b_category2"="Any warnings"
                   ))
```

## Culled population

Category AEP+ has a notable outlier, "better-enums".
We will remove it from the sample population to see how this affects the analysis results. 

```{r cyc-prepare-data-without-outliers, message=FALSE}
df_culled <- df
df_culled <- df_culled[df_culled$name != "better-enums",]

data_culled <- create_cyclomatic_complexity_data(df_culled)
data_culled$category <- factor(df_culled$category)

cyc7_culled <- fit_model_cyc7("cyc7_no_outliers", data_culled)
```

### Estimates {.tabset}

#### Standardized

```{r cyc-culled-standardized-estimates, message=FALSE}
summary(cyc7_culled)
```

#### Natural scale

```{r cyc-culled-natural-scale-estimates, message=FALSE}
print(create_natural_scale_summary_df(cyc7_culled,
                                      df_culled$cyclomatic_complexity_kloc))
```

### Plots {.tabset}

### Major category intervals

```{r cyc-culled-major-category-intervals, fig.height=2.8, message=FALSE}
plot_intervals_of_major_categories(cyc7_culled) +
  labs(title="CYC7 (culled)", x="Cyclomatic complexity per KLOC (std)")
```

### Binary warning usage

```{r cyc-culled-binary-warning-usage, fig.height=2, message=FALSE}
data_binary_factor_culled <- create_cyclomatic_complexity_data(df_culled)
data_binary_factor_culled$category <- factor(uses_warnings(df_culled$category))

cyc7_binary_factor_culled <-
  fit_model_cyc7("cyc7_no_outliers_factor_uses_warnings", 
                 data_binary_factor_culled)

plot_intervals_of_categories(cyc7_binary_factor_culled) +
  labs(title="CYC7 (culled)", x="Cyclomatic complexity per KLOC (std)") +
  scale_y_discrete(labels=c(
                     "b_category1"="No warnings",
                     "b_category2"="Any warnings"
                   ))
```
