---
title: Technical debt ratio
---

```{r tdr-include-setup, include=FALSE}
source(knitr::purl('setup.Rmd', output=tempfile()))
```

This section covers the analysis concerning the _technical debt ratio_ metric.

## Data preparation

```{r tdr-data-preparation, message=FALSE}
model_seed <- 20230302 # 2023 March 2

create_tech_debt_data <- function(data) {
  list(
       tech_debt_std=scale(data$sqale_debt_ratio),
    contributors_std=scale(data$contributors),
           stars_std=scale(data$stars),
             age_std=scale(get_age(data$created_at)),
           files_std=scale(data$files),
            size_std=scale(data$loc)
  )
}

data <- create_tech_debt_data(df)
data$category <- factor(df$category)
```

## Candidate models {.tabset}

### TD1

```{r tdr-model-1, message=FALSE}
fit_model_td1 <- function(name, data) {
  fit_model(name,
            "tech_debt_std ~ 0 + category",
            c(
              prior(normal(0, 1), class=b),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

td1 <- fit_model_td1("td1", data)
```

### TD2

```{r tdr-model-2, message=FALSE}
fit_model_td2 <- function(name, data) {
  fit_model(name,
            "tech_debt_std ~ 0 + category + files_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="files_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

td2 <- fit_model_td2("td2", data)
```

### TD3

```{r tdr-model-3, message=FALSE}
fit_model_td3 <- function(name, data) {
  fit_model(name,
            "tech_debt_std ~ 0 + category + contributors_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="contributors_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

td3 <- fit_model_td3("td3", data)
```

### TD4

```{r tdr-model-4, message=FALSE}
fit_model_td4 <- function(name, data) {
  fit_model(name,
            "tech_debt_std ~ 0 + category + size_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="size_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

td4 <- fit_model_td4("td4", data)
```

### TD5

```{r tdr-model-5, message=FALSE}
fit_model_td5 <- function(name, data) {
  fit_model(name,
            "tech_debt_std ~ 0 + category + age_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="age_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

td5 <- fit_model_td5("td5", data)
```

### TD6

```{r tdr-model-6, message=FALSE}
fit_model_td6 <- function(name, data) {
  fit_model(name,
            "tech_debt_std ~ 0 + category + stars_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="stars_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

td6 <- fit_model_td6("td6", data)
```

### TD7

```{r tdr-model-7, message=FALSE}
fit_model_td7 <- function(name, data) {
  fit_model(name,
            "tech_debt_std ~ 0 + category + contributors_std + files_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="contributors_std"),
              prior(normal(0, 1), class=b, coef="files_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

td7 <- fit_model_td7("td7", data)
```

### TD8

```{r tdr-model-8, message=FALSE}
fit_model_td8 <- function(name, data) {
  fit_model(name,
            "tech_debt_std ~ 0 + category + contributors_std + stars_std",
            c(
              prior(normal(0, 1), class=b),
              prior(normal(0, 1), class=b, coef="contributors_std"),
              prior(normal(0, 1), class=b, coef="stars_std"),
              prior(exponential(1), class=sigma)
            ),
            data,
            model_seed)
}

td8 <- fit_model_td8("td8", data)
```

## Model comparison

```{r tdr-model-comparison, message=FALSE}
td1 <- add_criterion(td1, "loo", moment_match=TRUE)
td2 <- add_criterion(td2, "loo", moment_match=TRUE)
td3 <- add_criterion(td3, "loo", moment_match=TRUE)
td4 <- add_criterion(td4, "loo", moment_match=TRUE)
td5 <- add_criterion(td5, "loo", moment_match=TRUE)
td6 <- add_criterion(td6, "loo", moment_match=TRUE)
td7 <- add_criterion(td7, "loo", moment_match=TRUE)
td8 <- add_criterion(td8, "loo", moment_match=TRUE)

print(loo_compare(td1, td2, td3, td4, td5, td6, td7, td8, criterion="loo"),
      simplify=FALSE)
```

## Diagnostics {.tabset}

### Prior predictive check

```{r tdr-ppc, message=FALSE}
plot_priors(td1, "Technical debt ratio (std)")
```

### MCMC chain trace plots

```{r tdr-caterpillars, message=FALSE}
plot_caterpillar(td1, "b_category1", "N")
plot_caterpillar(td1, "b_category2", "A")
plot_caterpillar(td1, "b_category3", "AE")
plot_caterpillar(td1, "b_category5", "AEP")
plot_caterpillar(td1, "b_category7", "AEP+")
```

## Full population

### Estimates {.tabset}

#### Standardized

```{r tdr-full-standardized-estimates, message=FALSE}
summary(td1)
```

#### Natural scale

```{r tdr-full-natural-scale-estimates, message=FALSE}
print(create_natural_scale_summary_df(td1, df$sqale_debt_ratio))
```

### Plots {.tabset}

#### Major category intervals

```{r tdr-full-major-category-intervals, fig.height=2.8, message=FALSE}
plot_intervals_of_major_categories(td1) +
  labs(title="TD1", x="Technical debt ratio (std)")
```

#### Binary warning usage

```{r tdr-full-binary-warning-usage, fig.height=2, message=FALSE}
data_binary_factor <- create_tech_debt_data(df)
data_binary_factor$category <- factor(uses_warnings(df$category))

td1_binary_factor <- fit_model_td1("td1_factor_uses_warnings", data_binary_factor)

plot_intervals_of_categories(td1_binary_factor) + 
  labs(title="TD1", x="Technical debt ratio (std)") +
  scale_y_discrete(labels=c(
                     "b_category1"="No warnings",
                     "b_category2"="Any warnings"
                  ))
```

## Culled population

```{r tdr-culled-data-preparation, fig.height=2.8, message=FALSE}
df_culled <- df
df_culled <- df_culled[df_culled$name != "clipp",]

data_culled <- create_tech_debt_data(df_culled)
data_culled$category <- factor(df_culled$category)

td1_culled <- fit_model_td1("td1_no_outliers", data_culled)
```

### Estimates {.tabset}

#### Standardized

```{r tdr-culled-standardized-estimates, message=FALSE}
summary(td1_culled)
```

#### Natural scale

```{r tdr-culled-natural-scale-estimates, message=FALSE}
print(create_natural_scale_summary_df(td1_culled, df_culled$sqale_debt_ratio))
```

### Plots {.tabset}

#### Major category intervals

```{r tdr-culled-major-category-intervals, fig.height=2.8, message=FALSE}
plot_intervals_of_major_categories(td1_culled) + 
  labs(title="TD1 (culled)", x="Technical debt ratio (std)")
```

#### Binary warning usage

```{r tdr-culled-binary-warning-usage, fig.height=2, message=FALSE}
data_binary_factor_culled <- create_tech_debt_data(df_culled)
data_binary_factor_culled$category <- factor(uses_warnings(df_culled$category))

td1_binary_factor_culled <-
  fit_model_td1("td1_data_no_outliers_factor_uses_warnings",
                data_binary_factor_culled)

plot_intervals_of_categories(td1_binary_factor_culled) +
   labs(title="TD1 (culled)", x="Technical debt ratio (std)") +
   scale_y_discrete(labels=c(
                     "b_category1"="No warnings",
                     "b_category2"="Any warnings"
                   ))
```
