# Setup

```{r, include=FALSE}
options(knitr.duplicate.label = "allow")
```

This page outlines the steps required to prepare the R environment.
First we need to load the required packages.

```{r setup-load-packages, message=FALSE}
# Bayesian data analysis tools
library(brms)

# Miscellaneous plotting
library(ggplot2)
library(ggpattern)
```

Then we will define some utility functions that are commonly used.

```{r setup-load-utility-functions}
# Returns the number of days between a date and the analysis date.
# We use February 28 as the reference date, otherwise it changes from day-to-day.
get_age <- function (yymmdd) {
  as.integer(difftime("2023-02-28", as.Date(yymmdd, "%Y-%m-%d"), units="days"))
}

# Adds columns for each metric that are relative to 1,000 lines-of-code.
add_scaled_metrics <- function(data) {
  data$bugs_kloc <- (data$bugs / data$loc) * 1000
  data$code_smells_kloc <- (data$code_smells / data$loc) * 1000
  data$critical_violations_kloc <- (data$critical_violations / data$loc) * 1000
  data$major_violations_kloc <- (data$major_violations / data$loc) * 1000
  data$minor_violations_kloc <- (data$minor_violations / data$loc) * 1000
  data$security_hotspots_kloc <- (data$security_hotspots / data$loc) * 1000
  data$vulnerabilities_kloc <- (data$vulnerabilities / data$loc) * 1000
  data$duplicated_lines_kloc <- (data$duplicated_lines / data$loc) * 1000
  data$cyclomatic_complexity_kloc <- (data$cyclomatic_complexity / data$loc) * 1000
  data$cognitive_complexity_kloc <- (data$cognitive_complexity / data$loc) * 1000
  data
}

# Returns 2 for projects that use warnings, 1 otherwise.
uses_warnings <- function(categories) {
  ifelse(categories %in% 2:11, 2, 1)
}

# Utility function used to fit brms models
fit_model <- function(name, formula, priors, data, seed) {
  brm(
    formula,
    family=gaussian,
    prior=priors,
    data=data,
    iter=10000,
    warmup=5000,
    chains=4,
    cores=4,
    file=paste("fits", name, sep="/"),
    file_refit="on_change",
    save_pars=save_pars(all=TRUE),
    seed=seed
  )
}

# Converts a standardized value to the natural scale, given the source data.
restore_from_std <- function(value, data) {
  value * sd(data) + mean(data)
}

# Plots credible intervals for all categories based on a fit model
plot_intervals_of_categories <- function(model) {
  mcmc_plot(model,
            type="intervals",
            variable="b_category",
            regex=TRUE,
            point_size=2,
            prob=0.5,        # Inner region density mass percentage
            prob_outer=0.89  # Outer region density mass percentage
  ) +
    theme_default(base_family="sans", base_size=14)
}

# Plots credible intervals for the major categories based on a fit model
plot_intervals_of_major_categories <- function(fit) {
  mcmc_plot(fit,
            type="intervals",
            variable=c(
              "b_category1",  # N
              "b_category2",  # A
              "b_category3",  # AE
              "b_category5",  # AEP
              "b_category7"   # AEP+
            ),
            regex=FALSE,
            point_size=2,
            prob=0.5,        # Inner region density mass percentage
            prob_outer=0.89  # Outer region density mass percentage
  ) +
    theme_default(base_family="sans", base_size=14) + 
    scale_y_discrete(labels=c(
                       "b_category1"="N",
                       "b_category2"="A",
                       "b_category3"="AE",
                       "b_category5"="AEP",
                       "b_category7"="AEP+"
                     ))
}
```

The data we need has been prepared in a separate CSV file, all we need to do is load it. However, we will need to create some new columns with metric values that are scaled according to the lines of code in each sample. This is so that we can compare metrics from projects of different sizes.

```{r setup-load-data-frame}
df <- read.csv("data/data_frame.csv")

# Include columns with metrics scaled per KLOC
df <- add_scaled_metrics(df)
```

We will also define some useful constants.

```{r setup-define-constants}
major_categories = c(1, 2, 3, 5, 7)
```

The following provides information about the used R session.

```{r setup-print-env-info}
sessionInfo()
```
